---
title: "Shiny"
author: "Samyuktha"
date: "2024-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Shiny is a web development package in R that allows you to turn your analyses into interactive web applications without needing to know HTML, CSS, or JavaScript. Shiny applications are easy to write and are built with R. 

## Installation and Getting Started

To get started, you need to install the `shiny` package. You can do this by running the following command:

```{r}
#install.packages("shiny")
```

Once you have installed the `shiny` package, you can create a new Shiny application by creating a new project in RStudio and switching to a new directory. You can then select "Shiny Application". It is recommended to create a new folder for each Shiny application you create. The project will consist of an Rproj file and an app.R file. The app.R file is where you will write your Shiny code.

The app.R file will contain an example Shiny application that you can run to see how it works. You can run the application by clicking on the "Run App" button in RStudio. 

## Components of a Shiny Application

The code consists of three main parts: the user interface (UI), the server and the function that connects the UI and the server. The UI is where you define the layout of the application, like the sidebar and the main panel, including the input controls and the output elements. The server is where you define the logic of the application, including how the inputs are processed and how the outputs are generated. 

## Examples

### Generating a histogram of random numbers

```{r}
library(shiny)

ui <- fluidPage(
  titlePanel("Random Number Generator"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("n", "Number of observations:", min = 1, max = 1000, value = 500)
    ),
    mainPanel(
      plotOutput("histogram")
    )
  )
)
  
server <- function(input, output) {
  output$histogram <- renderPlot({
    hist(rnorm(input$n))
  })
}

shinyApp(ui = ui, server = server)
```
This code creates a Shiny application with a slider input control that allows the user to select the number of observations to generate. The application then generates a histogram of random numbers based on the user's input.

You can customize the appearance and behavior of your Shiny application by adding additional input controls, output elements, and interactive features. Shiny provides a wide range of functions and widgets that you can use to create interactive web applications that are tailored to your specific needs.

### Simulating the Lotka-Volterra Predator-Prey Dynamics

This code creates a Shiny application that simulates the Lotka-Volterra predator-prey dynamics. The user can adjust the parameters of the system and observe how the populations of the prey and predator change over time. The application uses the `deSolve` package to solve the system of ordinary differential equations that describe the dynamics of the system.


```{r}
library(deSolve)
library(shiny)

# Defining the differential equations
lotka_volterra <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    dxdt = alpha * x - beta * x * y
    dydt = delta * x * y - gamma * y
    return(list(c(dxdt, dydt)))
  })
}

# Defining the UI
ui <- fluidPage(
  titlePanel("Lotka-Volterra Predator-Prey Dynamics"),

  sidebarLayout(
    sidebarPanel(
      numericInput("alpha", "Prey growth rate (alpha):", 1),
      numericInput("beta", "Predator-prey interaction rate (beta):", 0.1),
      numericInput("gamma", "Predator death rate (gamma):", 1),
      numericInput("delta", "Predator reproduction rate (delta):", 0.1),
      numericInput("x0", "Initial prey population x(0):", 40),
      numericInput("y0", "Initial predator population y(0):", 9),
      numericInput("time", "Time range:", 100, min = 1, step = 1)
    ),

    mainPanel(
      plotOutput("solutionPlot")
    )
  )
)

# Defining the server
server <- function(input, output) {

  # Reactive expression to solve the Lotka-Volterra system
  solve_system <- reactive({
    # Parameters
    parameters <- c(alpha = input$alpha, beta = input$beta, gamma = input$gamma, delta = input$delta)

    # Initial state
    state <- c(x = input$x0, y = input$y0)

    # Time 
    time <- seq(0, input$time, by = 0.1)

    # Solving the system of ODEs
    out <- ode(y = state, times = time, func = lotka_volterra, parms = parameters)

    # Converting the output to a data frame for plotting
    data.frame(Time = out[, 1], Prey = out[, 2], Predator = out[, 3])
  })

  # Plotting the solution
  output$solutionPlot <- renderPlot({
    result <- solve_system()

    plot(result$Time, result$Prey, type = "l", col = "blue", lwd = 2,
         xlab = "Time", ylab = "Population", ylim = range(c(result$Prey, result$Predator)))
    lines(result$Time, result$Predator, col = "red", lwd = 2)
    legend("topright", legend = c("Prey (x)", "Predator (y)"), col = c("blue", "red"), lwd = 2)
  })
}

# Running the application
shinyApp(ui = ui, server = server)

```

## Deployment

Shiny applications can be deployed to the web using the `shinyapps.io` service, which allows you to host your applications online and share them with others. Here is an example of a Shiny application that has been deployed to `shinyapps.io`: [https://samyuktha-ramadurai.shinyapps.io/Lotka-Volterra/](https://samyuktha-ramadurai.shinyapps.io/Lotka-Volterra/)

## Resources

https://shiny.posit.co/r/articles/#start  <br />
[Tutorials](https://shiny.posit.co/r/getstarted/shiny-basics/lesson1/)  <br />
[Cheat Sheet](https://shiny.posit.co/r/articles/start/cheatsheet/) <br />
[Posit's YouTube Demos and Tutorials](https://www.youtube.com/playlist?list=PL9HYL-VRX0oTOK4cpbCbRk15K2roEgzVW) <br />
[Wickham’s book](https://mastering-shiny.org/)


